---
title: 'STAT 341: Tutorial 2 -- Pokémon Analysis (Part 1)'
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "Friday January 17, 2020"
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we begin to take a look at the `pokemon.csv` dataset introduced in the first lecture. This exercise is intended to provide additional exposure to R and Markdown.

**Load in the data and summarize it**
```{r, echo=TRUE, eval=TRUE}
poke <- read.csv("/Users/nstevens/Dropbox/Teaching/STAT_341/Tutorials/Tutorial 2/pokemon.csv", 
                 header = TRUE)
```

Now that we've loaded the data, let's look at a summary of it:
```{r, echo=FALSE, eval=TRUE}
summary(poke)
```

For purposes of the tutorial, let's examine the `base_happiness`, `speed`, `height_m` and `weight_kg` variates more closely. Here are the values of these variates for the first 15 Pokémon:
```{r}
head(x = poke[,c(4, 8, 9, 10)], n = 15)
```

Let's construct a $4 \times 4$ matrix of plots that visually summarize these:
```{r, echo=TRUE, eval=TRUE, fig.align='center', fig.height=8, fig.width=8}
poke_df_for_plot <- poke[,c(4, 8, 9, 10)]
header <- names(poke_df_for_plot)
par(mfrow = c(4,4))
for(i in 1:4){
  for(j in 1:4){
    if(i == j){
      hist(poke_df_for_plot[,i], main = "", xlab = header[i], col = "dodgerblue3")
    }else{
      plot(x = poke_df_for_plot[,i], y = poke_df_for_plot[,j],
           xlab = header[i], ylab = header[j],
           pch = 16, col = adjustcolor(col = "dodgerblue4", alpha.f = 0.3))
    }
  }
}
```

Since `base_happiness` takes on just a handful of integer values, let's _jitter_ this variate to try and make the plots more informative.
```{r, fig.align='center', fig.height=8, fig.width=8}
poke_df_for_plot$base_happiness <- jitter(poke_df_for_plot$base_happiness, factor = 3)
par(mfrow = c(4,4))
for(i in 1:4){
  for(j in 1:4){
    if(i == j){
      hist(poke_df_for_plot[,i], main = "", xlab = header[i], col = "dodgerblue3")
    }else{
      plot(x = poke_df_for_plot[,i], y = poke_df_for_plot[,j],
           xlab = header[i], ylab = header[j],
           pch = 16, col = adjustcolor(col = "dodgerblue4", alpha.f = 0.3))
    }
  }
}
```

Let's also calculate the correlations between these variables:
```{r, echo=FALSE, eval=TRUE}
cor(poke[,c(4, 8, 9, 10)])
```

**Questions:**

1. Do faster Pokémon tend to be happier?
2. Do taller Pokémon tend to be happier?
3. Do heavier Pokémon tend to be happier?

**Answers:**

1. No
2. No
3. No

**Who is the tallest Pokémon?**
```{r}
# this is the maximum height:
max(poke$height_m)
# this is the row in the dataset in which the maximum height is found:
which(poke$height_m == max(poke$height_m))
# this is the corresponding Pokémon:
poke$name[which(poke$height_m == max(poke$height_m))]
```

Note that in the code above, rather than the call
```{r, eval=FALSE}
which(poke$height_m == max(poke$height_m))
```

we could have equivalently used
```{r, eval=FALSE}
which.max(poke$height_m)
```

**Which are the 10 heaviest Pokémon?**
```{r}
# these are the weights sorted from largest to smallest
sort(x = poke$weight_kg, decreasing = TRUE)
# these are the 10 largest recorded weights
sort(x = poke$weight_kg, decreasing = TRUE)[1:10]
# these are the rows containing the 10 largest recorded weights
which(poke$weight_kg >= sort(x = poke$weight_kg, decreasing = TRUE)[10])
# these are the corresponding Pokémon
poke$name[which(poke$weight_kg >= sort(x = poke$weight_kg, decreasing = TRUE)[10])]
```

**Construct a histogram of `hp` with the mean identified**
```{r, fig.align='center'}
hist(x = poke$hp, col = "mistyrose", main = "Histogram of HP", xlab = "HP")
abline(v = mean(poke$hp), col = "navyblue", lwd = 2)
text(x = 70, y = 200, labels = paste("Mean = ", round(mean(poke$hp), 2)),
     col = "navyblue", pos = 4)
```

**Plot the influence ($\Delta$) of each Pokémon on the average `hp`**
```{r, fig.align='center'}
y <- poke$hp
delta <- (y-mean(y))/(length(y)-1)
par(mfrow = c(1,2))
plot(delta,
     main = "Influence for Average HP", ylab = bquote(Delta),
     pch = 16, col = adjustcolor("mediumvioletred", 0.3))
plot(y, delta,
     main = "Influence vs. HP", ylab = bquote(Delta), xlab = "HP",
     pch = 16, col = adjustcolor("mediumvioletred", 0.3))
```

**Which two Pokémon have the largest influence?**
```{r}
poke$name[which(delta > 0.2)]
```
**Which two Pokémon have the highest HP?**
```{r}
poke$name[order(poke$hp, decreasing = TRUE)][1:2]
```